\newgeometry{
	marginratio={1:1, },
	bottom=0.75in,
	top=0.8in,
	textwidth=2.3\alphabet,
}
\section*{References}

\begin{itemize}[label={}, leftmargin=*, itemsep={0.5\baselineskip plus 2pt}]
\citem{bayes}{} T.~Bayes (1763), Phil.~Trans.~Roy.~Soc.~330-418.~Reprint, with biographical note by G.~A.~Barnard, in Biometrika \emph{45}, 293 (1958) and in \emph{Studies in the History of Statistics and Probability}, E.~S.~Pearson and M.~G.~Kendall, editors, C.~R.~Griffin and Co.~Ltd., London (1970). Also reprinted in \emph{Two Papers by Bayes with Commentaries}, W.~E.~Deming, editor, Hafner Publishing Co., New York (1963).

\citem{bell} E.~T.~Bell (1937), \emph{Men of Mathematics}, Simon and Schuster, N.Y.

\citem{boltzmann} L.~Boltzmann (1871), Wien.~Ber.~\emph{63}, 397, 679, 712.

\citem{boole} G.~Boole (1854), \emph{The Laws of Thought}, Reprinted by Dover Publications, Inc., New York.

\citem{cox} R.~T.~Cox (1946), Am.~J.~Phys.~\emph{17}, 1. See also \emph{The Algebra of Probable Inference}, Johns Hopkins University Press (1961); reviewed by E.~T.~Jaynes in Am.~J.~Phys.~\emph{31}, 66 (1963).

\citem{czuber} E.~Czuber (1908), \emph{Wahrscheinlichkeitsrechnung und Ihre Anwendung auf Fehlerausgleichung}, Teubner, Berlin. Two Volumes.

\citem{ehrenfest} P.~and T.~Ehrenfest (1912), Encykl.~Math.~Wiss. English translation by M.~J.~Moravcsik, \emph{The Conceptual Foundations of the Statistical Approach in Mechanics}. Cornell University Press, Ithaca, N. Y. (1959).

\citem{ellis} P.~L.~Ellis (1842), ``On the Foundations of the Theory of Probability'', Camb. Phil. Soc. vol.~viii. The quotation given by Rowlinson actually appears in a later work: Phil.~Mag.~vol.~xxxvii (1850). Both are reprinted in his \emph{Mathematical and other Writings} (1863).

\citem{friedman} K.~Friedman and A.~Shimony (1971), J.~Stat.~Phys.~\emph{3}, 381. See also M.~Tribus and H.~Motroni, \emph{ibid.~4}, 227; A.~Hobson, \emph{ibid.~6}, 189; D.~Gage and D.~Hestenes, \emph{ibid.~7}, 89; A.~Shimony, \emph{ibid.~9}, 187; K.~Friedman, \emph{ibid.~9}, 265.

\citem{gibbs} J.~W.~Gibbs (1902), \emph{Elementary Principles in Statistical Mechanics}, Reprinted in Collected works and commentary, Yale University Press (1936), and by Dover Publications, Inc.~(1960).

\citem{good} I.~J.~Good (1950), \emph{Probability and the Weighing of Evidence}, C.~Griffin and Co., London.

\citem{heath} D.~Heath and Wm.~Sudderth (1976), ``de Finetti's Theorem on Exchangeable Variables,'' The American Statistician, \emph{30}, 188.

\citem{haar} D.~ter~Haar (1954), \emph{Elements of Statistical Mechanics}, Rinehart and Co., New York.

\citem{heims} S.~P.~Heims and E.~T.~Jaynes (1962), Rev.~Mod.~Phys.~\emph{34}, 143.

\citem{huzurbazar} V.~S.~Huzurbazar (1976), \emph{Sufficient Statistics} (Volume 19 in Statistics Textbooks and Monographs Series) Marcel Dekker, Inc., New York.

\citem{jaynes57} E.~T.~Jaynes (1957a,b), Phys.~Rev.~\emph{106}, 620; \emph{108}, 171.

\citem{jaynes63a} E.~T.~Jaynes (1963a), ``New Engineering Applications of Information Theory,'' in Proceedings of the \emph{First Symposium on Engineering Applications of Random Function Theory and Probability}, J.~L.~Bogdanoff and F.~Kozin, editors; J.~Wiley and Sons, New York, pp.~163-203.

\citem{jaynes63b} E.~T.~Jaynes (1963b), ``Information Theory and Statistical Mechanics,'' in \emph{Statistical Physics} (1962 Brandeis Lectures) K.~W.~Ford, ed.~W.~A.~Benjamin, Inc., New York, pp.~181-218.

\citem{jaynes65} E.~T.~Jaynes (1965), ``Gibbs vs.~Boltzmann Entropies,'' Am.~J.~Phys.~\emph{33}, 391.

\citem{jaynes67} E.~T.~Jaynes (1967), ``Foundations of Probability Theory and Statistical Mechanics,'' in \emph{Delaware Seminar in the Foundations of Physics}, M.~Bunge, ed., Springer-Verlag, Berlin; pp.~77~101.

\citem{jaynes68} E.~T.~Jaynes, (1968), ``Prior Probabilities,'' IEEE Trans.~on Systems Science and Cybernetics, \emph{SSC-4}, pp.~227-241. Reprinted \emph{in Concepts and Applications of Modern Decision Models}, V.~M.~Rao Tummala and R.~C.~Henshaw, eds., (Michigan State University Business Studies Series, 1976).

\citem{jaynes71} E.~T.~Jaynes (1971), ``Violation of Boltzmann's H-theorem in Real Gases,'' Phys.~Rev.~A \emph{4}, 747.

\citem{jaynes73} E.~T.~Jaynes (1973), ``The Well-Posed Problem,'' Found.~Phys.~\emph{3}, 477.

\citem{jaynes76} E.~T.~Jaynes (1976), ``Confidence Intervals vs.~Bayesian Intervals,'' in \emph{Foundations of Probability Theory, Statistical Inference, and Statistical Theories of Science}, W.~L.~Harper and C.~A.~Hooker, eds., D.~Reidel Publishing Co., Dordrecht-Holland, pp.~175-257.

\citem{jaynes78} E.~T.~Jaynes (1978), ``Electrodynamics Today,'' in \emph{Proceedings of the Fourth Rochester Conference on Coherence and Quantum Optics}, L.~Mandel and E.~Wolf, eds., Plenum Press, New York.

\citem{jaynes79} E.~T.~Jaynes (1979), ``Marginalization and Prior Probabilities,'' in \emph{Studies in Bayesian Econometrics and Statistics}, A.~Zellner, ed., North-Holland Publishing Co., Amsterdam (in press).

\citem{jeffreys} H.~Jeffreys (1939), \emph{Theory of Probability}, Oxford University Press.

\citem{keynes} J.~M.~Keynes (1921), \emph{A Treatise on Probability}, MacMillan and Co., London.

\citem{kubo57} R.~Kubo (1957), J.~Phys.~Soc.~(Japan) \emph{12}, 570.

\citem{kubo58} R.~Kubo (1958), ``Some Aspects of the Statistical-Mechanical Theory of Irreversible Processes,'' in \emph{Lectures in Theoretical Physics}, Vol.~1, W.~E.~Brittin and L.~G.~Dunham, eds, Interscience Publishers, Inc., New York; pp.~120-203.

\citem{kullback} S.~Kullback (1959), \emph{Information Theory and Statistics}, J.~Wiley and Sons, Inc.

\citem{lindhard} J.~Lindhard (1974), ``On the Theory of Measurement and Its Consequences in Statistical Dynamics,'' Kgl.~Danske.~Vid. Selskab Mat. \mbox{-Fys}.~Medd. \emph{29}, 1.

\citem{maxwell59} J.~C.~Maxwell (1859), ``Illustrations of the Dynamical Theory of Gases,'' \emph{Collected Works}, W.~D.~Niven, ed., London (1890), Vol.~I, pp.~377-409.

\citem{maxwell76} J.~C.~Maxwell (1876), ``Qn Boltzmann's Theorem on the Average Distribution of Energy in a System of Material Points,'' \emph{ibid.} Vol.~II, pp.~713-741.

\citem{middleton} D.~Middleton (1960), \emph{Introduction to Statistical Communication Theory}, Mc\-Graw-Hill Book Company, Inc., New York, pp.~1082-1102.

\citem{mori} H.~Mori, I.~Oppenheim, and J.~Ross (1962), ``Some Topics in Quantum Statistics,'' in \emph{Studies in Statistical Mechanics}, Volume~1, J.~de~Boer and G.~E.~Uhlenbeck, eds., North-Holland Publishing Company, Amsterdam; pp.~213-298.

\citem{polya} G.~Polya (1954), \emph{Mathematics and Plausible Reasoning}. Princeton University Press, two volumes.

\citem{robertson} B.~Robertson (1964), Stanford University, Ph.D.~Thesis.

\citem{rowlinson} J.~S.~Rowlinson (1970), ``Probability, Information, and Entropy,'' Nature, 225, 1196.

\citem{savage} L.~J.~Savage (1954), \emph{The Foundations of Statistics}, John Wiley and Sons, Inc,, Revised Edition by Dover Publications, Inc., 1972.

\citem{schrodinger} E.~Schrodinger (1948), \emph{Statistical Thermodynamics}, Cambridge University Press.

\citem{shannon} C.~Shannon (1948), Bell System Tech.~J. \emph{27}, 379, 623. Reprinted in C.~E.~Shannon and W.~Weaver, \emph{The Mathematical Theory of Communication}, University of Illinois Press, Urbana (1949).

\citem{truesdell} C.~Truesdell (1960), ``Ergodic Theory in Classical Statistical Mechanics,'' in \emph{Ergodic Theories}, Proceedings of the International School of Physics ``Enrico Fermi,'' P.~Caldirola, ed., Academic Press, New York, pp.~21-56.

\citem{venn} J.~Venn (1866), \emph{The Logic of Chance}

\citem{mises} R.~von Mises (1928), \emph{Probability, Statistics and Truth}, J.~Springer. Revised English Edition, H.~Geiringer, ed., George Allen and Unwin Ltd., London (1957).

\citem{wald} A.~Wald (1950), \emph{Statistical Decision Functions}, John Wiley and Sons, Inc., New York.

\citem{watson} J.~D.~Watson (1965), Molecular Biology of the Gene, W.~A. Benjamin, Inc., New York.~See particularly pp.~73-100.

\citem{wiener} N.~Wiener (1949), \emph{Extrapolation, Interpolation, and Smoothing of Stationary Time Series}, Technology Press, M.I.T.

\citem{wilks} S.~S.~Wilks (1961), Foreword to Geralomo Cardano, \emph{The Book on Games of Chance}, translated by S.~H.~Gould.~Holt, Rinehart, and Winston, New York.

\citem{zwanzig} R.~Zwanzig (1960), J.~Chem.~Phys.~\emph{33}, 1338; Phys.~Rev.~\emph{124}, 983 (1961).
\end{itemize}

\restoregeometry